---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# faircfint

<!-- badges: start -->
<!-- badges: end -->

The goal of faircfint is to facilitate estimation and inference for intersectional, counterfactual unfairness metrics. The metrics and techniques are described in ["An intersectional framework for counterfactual fairness in risk prediction"](https://arxiv.org/abs/2210.01194) (2022).

## Installation

You can install the development version of faircfint from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("swastvedt/faircfint")
```

## Usage

The main function, analysis_estimation, calculates the intersectional, counterfactual unfairness metrics from data (which must include risk predictions).

```{r}
library(faircfint)

# Build propensity score model
pi_model_ex <- glm(as.factor(as.character(D)) ~ A1*A2 + A1 + A2 + S_prob + X.1 + X.2 + X.3 + X.4, 
                   data = faircfint::data_est, family = "binomial")
## Propensity score model formula
pi_formula_ex <- as.formula(pi_model_ex$formula)

# Estimate unfairness metrics, including bootstrap and null distribution
results <- analysis_estimation(data = faircfint::data_est, cutoff = 0.5, 
                               pi_model = pi_model_ex, pi_model_type = "glm", 
                               f_lasso = pi_formula_ex, 
                               xvars = c("X.1", "X.2", "X.3", "X.4"), gen_null = T,
                               bootstrap = 'rescaled')

# Plot results
results_plots <- get_plots(results, sampsize = 5000, alpha = 0.5, m_factor = 0.75, 
                           A1_length = 2, A2_length = 2, 
                           plot_labels = c("Group 1", "Group 2", "Group 3", "Group 4"),
                           plot_values = c(15,16,17,18), 
                           plot_colors = c("black", "gray", "red", "dodgerblue"))

# Example plot of negative (cFNR) versions of metrics
#results_plots$cfnr
```


<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>. -->

