---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# faircfint

<!-- badges: start -->
<!-- badges: end -->

The goal of faircfint is to facilitate estimation and inference for intersectional, counterfactual unfairness metrics. The metrics and techniques are described in ["An intersectional framework for counterfactual fairness in risk prediction"](https://doi.org/10.1093/biostatistics/kxad021) and ["Counterfactual fairness for small subgroups"](https://arxiv.org/abs/2310.19988).

## Installation

You can install the development version of faircfint from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("swastvedt/faircfint")
```

## Usage

The main function, analysis_estimation, calculates the intersectional, counterfactual unfairness metrics from data (which must include risk predictions). 

The 'standard' option for estimator_type uses the methods from ["An intersectional framework for counterfactual fairness in risk prediction"](https://doi.org/10.1093/biostatistics/kxad021). Options 'small_internal' and 'small_borrow' use the methods from ["Counterfactual fairness for small subgroups"](https://arxiv.org/abs/2310.19988).

```{r}
library(faircfint)

# Build propensity score model
pi_model_ex <- glm(as.factor(as.character(D)) ~ A1*A2 + A1 + A2 + S_prob + X.1 + X.2 + X.3 + X.4, 
                   data = faircfint::ex_data_estimation, family = "binomial")

# Estimate counterfactual error rates and unfairness metrics, including bootstrap and null distribution
results <- analysis_estimation(data = faircfint::ex_data_estimation, cutoff = 0.5, estimator_type = "standard",
                               pi_model = pi_model_ex, pi_model_type = "glm", pi_xvars = c("X.1", "X.2", "X.3", "X.4"),
                               gen_null = T, bootstrap = 'rescaled')

# Estimate metrics using small subgroup option with no data borrowing
pi_model_ex_small <- glm(as.factor(as.character(D)) ~ A1*A2 + A1 + A2 + S_prob + X.1 + X.2 + X.3 + X.4,
                         data = ex_data_small, family = "binomial")

results_smallsub <- analysis_estimation(ex_data_small, cutoff = 0.5, estimator_type = "small_internal",
                                     gen_null = T, R_null = 100, bootstrap = 'rescaled', B=100,
                                     pi_model = pi_model_ex_small, pi_model_type = "glm", pi_xvars = c("X.1", "X.2", "X.3", "X.4"),
                                     outcome_model_type = "glm", outcome_xvars = c("X_outcome.1", "X_outcome.2", "X_outcome.3", "X_outcome.4", "X_outcome.5", "X_outcome.6", "X_outcome.7", "X_outcome.8"),
                                     fit_method_int = "multinomial", nfolds = 5, pa_xvars_int = c("X_pa.1", "X_pa.2", "X_pa.3", "X_pa.4", "X_pa.5", "X_pa.6", "X_pa.7"))

# Estimate metrics using small subgroup option with data borrowing
## Optional fitting of external data model prior to running main function
pa_model_ext_ex <- get_pa_ext_small(ex_data_external, fit_method_ext = "multinomial", pa_xvars_ext = c("X_pa.1", "X_pa.2", "X_pa.3", "X_pa.4", "X_pa.5"))

results_borrow <- analysis_estimation(ex_data_small, cutoff = 0.5, estimator_type = "small_borrow",
                                         gen_null = F, R_null = 100, bootstrap = 'none', B=100,
                                         pi_model = pi_model_ex_small, pi_model_type = "glm", pi_xvars = c("X.1", "X.2", "X.3", "X.4"),
                                         outcome_model_type = "glm", outcome_xvars = c("X_outcome.1", "X_outcome.2", "X_outcome.3", "X_outcome.4", "X_outcome.5", "X_outcome.6", "X_outcome.7", "X_outcome.8"),
                                         fit_method_int = "multinomial", nfolds = 5, pa_xvars_int = c("X_pa.1", "X_pa.2", "X_pa.3", "X_pa.4", "X_pa.5", "X_pa.6", "X_pa.7"),
                                         data_external = ex_data_external, fit_method_ext = "multinomial", pa_xvars_ext = c("X_pa.1", "X_pa.2", "X_pa.3", "X_pa.4", "X_pa.5"), borrow_metric = "brier",
                                         pa_model_ext = pa_model_ext_ex)

# Plot results
results_plots <- get_plots(results, sampsize = 5000, alpha = 0.05, m_factor = 0.75)


# Example plot of negative (cFNR) versions of small subgroup metrics
#results_plots$metrics_pos
```

